%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  ch-magic.tex
%%
%%  Created: Fri Oct 10 14:24:37 1997
%%  Author.: Jose Carlos Gonzalez
%%  Notes..:
%%          
%%-------------------------------------------------------------------------
%% Filename: $RCSfile$
%% Revision: $Revision$
%% Date:     $Date$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{The \MAGIC Telescope}
\label{chapter:magic}

The \I{MAGIC} Telescope (\emph{Major Atmospheric Gamma-rays Imaging
  Cherenkov Telescope}) is a pro-ject devoted to the construction of a
17\u{m} diameter \Cherenkov telescope, with the most advanced
technology available in this field. Many of the new elements used and
developed especially for this project, among other technical
innovations, are: low weight, high reflectivity coated aluminum
mirrors; \I{active mirror control}, fast \I{Flash ADC}s, \I{optical
  fiber} links, fast slew for \I{GRB}s observation, multilevel
trigger, atmospheric monitoring, etc.

Nowadays $\gamma$-ray experiments on board of satellites can detect
photons with energies as high as $\sim$10\u{GeV}. Higher energies are
impossible to detect with such devices, due to their small detection
area.  On the other hand, the most advanced Cherenkov detectors
currently operatives have a detection threshold above 250--300\u{GeV}.
Below this values, their sensitivity drops down drastically. The main
goal of the \MAGIC Telescope (\MAGIC hereafter) is to cover this
energy range between 10\u{GeV} and 200\u{GeV}.
%
\energygapfig

The detector \I{EGRET} (\emph{Energetic Gamma Ray Experiment
  Telescope}), installed on board of the \emph{Compton Gamma-Ray
  Observatory} (\I{CGRO}), and with an upper detection threshold of
10\u{GeV}), has detected $\gamma$-rays from numerous sources.
However, from the set of observed extra-galactic
sources, most of sources with high %
%
$z$\footnote{The magnitude denoted by $z$ is the so called
  \emph{redshift}.  It gives the observed deviation in the wavelength
  from a wave coming from a source moving away from us.  Its
  calculation can be done using the expression
  $\Delta\lambda/\lambda=v/c=z$, where $v$ is the velocity of the
  source, and $c$ is the speed of light in vacuum.} %
%
have not been observed at energies above a few hundred \u{GeV} using
other detectors.  In addition, there are only three detected
\emph{\Isee{supernov{\ae}\ remnant}{SNR}} (\I{SNR} hereafter): the Crab
Nebula, Vela and \I{SN\,1006}, at energies of few \u{TeV}'s, despite
the assumption of being (the SNR) one of the main sources of cosmic
radiation.  In both cases (extra-galactic sources and supernov{\ae}) the
detection could be possible in the range 10--200\u{GeV}.  Therefore,
it is clear that this energy window is of extreme importance for the
$\gamma$-ray astronomy.

\CGROenergiesfig

\section{Physical goals}
%
We can say, there is a major physical goal for the \MAGIC Telescope,
namely, \emph{covering the unexplored energy range from 10\u{GeV} to
  100\u{GeV}}. This global aim is expressed, however, in a variety of
more specific targets, which now I will try to enumerate:

\subsubsection*{Active Galactic Nuclei (AGN)} 
%
About 1\% of the galaxies are found to be active. Active means, in this
context, that with their core is associated a highly luminous and
variable non-thermal emission. In this big class of objects we can
find the radio-galaxies, the Seyfert galaxies, quasars (quasi-stellar
radio objects, both radio-loud and radio-quiet), and the BL-Lacert{\ae}\ 
(BL-Lac) objects. A detailed classification of the AGN was shown in
Fig. \ref{fig:AGNclassification}.

%\AGNclassificationfig

Since the detection of several AGN by EGRET, the attention has been
focussed on the \emph{blazar} subclass. This subclass consists of the
\emph{BL Lac objects} and the \emph{Flat Spectrum Radio Quasars}
(FSRQs).  Both types of objects are characterized by flat
radio-spectrum, strong variability and polarization of the optical
emission.  The blazars are the main candidates for the high and very
high energy gamma-ray emission.

The main questions related to the AGN observations are:
%
\begin{itemize}
\item How do AGN form and evolve?
\item What do the jets (relativistic plasma outflows) from AGN consist
  of and how is this plasma accelerated?
\item Is there a \emph{cut-off} for the photon energy spectrum? How
  and where are these cut-offs caused?
\item What's the reason for the observed strong variability?
\end{itemize}

Blazar emission for X-ray selected BL Lac objects at lower energies
(up to about 1-100 keV) is almost certainly due to synchrotron
emission from a beam of highly relativistic electrons.  Above
300\u{GeV}, only three AGN have been detected so far (and one of them
only by the Whipple Telescope, in a short flare). But more important
than this: from these AGN, only one (Mkn421) has been detected by
EGRET (actually, there is a report of a weak detection of Mkn501).
Since EGRET detected several tens of AGN at energies below 30\u{GeV},
most of these AGN must have their cut-off at energies between 30 and
300\u{GeV}. There exist several scenarios with different explanations.
Under the assumption that the \emph{diffuse extra-galactic
  $\gamma$-background} might be due to the sum of the emission of
unresolved blazars, the spectrum of these AGN should extend almost
unchanged in slope up to 50--100\u{GeV}.

Moreover, there could be a whole sub-class of AGN with its second peak
in the spectral energy distribution above 10\u{GeV} (the first peak in
the spectrum is between UV to X-rays). These AGN would be difficult to
detect by satellite experiments. This makes the study of AGN in this
unexplored energy range especially interesting for modeling blazars.

\subsubsection*{Supernov{\ae}\ remnants (SNRs)}
% 
As we already mentioned, only three SNRs are related to $\gamma$-ray
emission: the Crab Nebula, Vela and SN1006. The production mechanism
for $\gamma$-rays is believed to be Inverse Compton Scattering (ICS)
of synchrotron and background field photons by VHE electrons, or
$\pi^0$ decay into $\gamma$s after the interaction of accelerated
hadrons with target matter external to the remnant. The main question
that link MAGIC with the study of SNRs are:
%
\begin{itemize}
\item Which  kind of particles  are responsible for  the production of
  high and very high energy $\gamma$s in SNRs?
\item Is  the mechanism  of production  of $\gamma$s dependent  on the
  type of SNR?
\item Are SNRs really responsible for the bulk of cosmic rays?
\end{itemize}

\subsubsection*{Pulsars} 
%
Pulsars are the expected energy sources in Plerion type supernova
remnants. Apart from the Crab pulsar, another two pulsars have been
detected at very high energies: the Vela pulsar (plerion) and
PSR\,1706-44. All the pulsars detected by EGRET seem to have quite
different properties. The most recent theoretical considerations
suggest that pulsed emission from isolated pulsars cannot be expected
beyond 6--20\u{GeV} in the \emph{polar-cap model}, or around
100\u{GeV} in the \emph{outer-cap model}.  This is already more than
enough to study the radio-pulsars in the MAGIC energy range.

Concerning to radio-quiet pulsars, only one has been so far detected
at high energies: the Geminga pulsar. If we assume this is the closest
pulsar of its type, one can estimate the number of such objects in our
galaxy to be around 1600. MAGIC will look for objects fainter than
Geminga. In this set are also the faintest unidentified EGRET sources.

\subsubsection*{Gamma Ray Bursts} 
%
A clear milestone in the current status of the Gamma Ray Astrophysics
is the nature of the GRBs. With MAGIC we will be able to measure the
power spectrum of GRBs. MAGIC will be able of fast positioning (from
one position to any other in less than 30 seconds).  This, together
with the help of fast timing and positioning information from future
detectors on board of satellites will allow us to record the bulk of
the flux emitted by GRBs ($\sim$1--100\u{s}, although it can be longer
--- for the GRB\,971208 the single peaked burst lasted around
800\u{sec}).

\subsubsection*{Unidentified EGRET sources} 
%
We have already mentioned earlier the importance of look for the
unidentified EGRET sources in the MAGIC energy range 10--200\u{GeV}.
About 40\% of these sources have hard power spectra which seem to
extend beyond 10\u{GeV}. The observation of these objects with MAGIC
will enhance the accuracy of the location from 0.1\deg to 0.02\deg,
and therefore the possibility of identification and study of these
sources.

\MAGICframefig

\section{Optical characteristics of the reflector}
%
The \MAGIC telescope will have a 17\u{m} diameter octagonal reflector.
The mount is alt-azimuthal. The reflector is supported by a 4 layers
space frame, and is fixed at two sides onto the azimuth drive by
stub-axes. This space frame is assembled from low-weight carbon
fiber-epoxy tubes and aluminum knots. The dish is mounted inside a
ring (the $\theta$-ring) which holds the camera at its upper apex,
while the lower section is equipped with a drive chain for altitude
movement (see Fig.\ref{fig:MAGICframe}).  The telescope has been
designed to have a f/D:1 optics, i.e. the global focal length of the
system is 17\u{m}.  The $\theta$-ring has an elliptical shape
(\emph{gothic arc}), in order to avoid the 30--32\u{m} diameter
circular ring needed to follow the f/D:1 optics.

\MAGICmirrorcontrol_wfig
%
The global shape of the dish is parabolic, and is tessellated: it is
composed of 972 small squared mirrors, $50\times 50 \u{cm}^2$ each.
The total mirrored surface is then $230\u{m}^2$. The curvature of
these small mirrors is spherical.  This design allows a small
\emph{spherical aberration} together with a small time spread of the
signal in the camera. Due to the relatively large increase in distance
of the mirrors to the focus of the telescope, their curvature radii
are taken to vary between 34\u{m} and 36\u{m}, approximately.  The
small square mirrors ($50\times 50 \u{cm}^2$) are distributed
following perpendicular rows and columns. They are grouped in sets of
four, forming an individual element of $1\times 1 \u{m}^2$.  Each of
these sets will be independently adjusted and focussed (\emph{active
  optics}).

The deformation of the frame will be monitorized by a laser beam
pointed to several semi-transparent silicon strip sensors, with an
overall resolution in the measurement of the deformations on the order
of few micrometers. The mirror elements, in groups of four, will be
focussed and aligned used an \emph{active mirror control
  system}\label{label:amcs}.  Briefly, a laser pointer located on each
panel of four mirrors will be pointed towards the center of the
camera, exactly to where the four mirrors are focused. The spot in the
camera plane will be checked using a video camera, and two stepping
motors, driven via a multiplexer by a computer-controlled drive
circuit, will adjust the position and orientation of the panel (see
Fig.\ref{fig:MAGICmirrorcontrol}).

\label{label:mirrorelement}%
The single mirror elements have different focal distances, and
therefore the surfaces have to be machined on a numerically controlled
machine. \MAGIC will use all-aluminum sandwich mirrors, similar to
those already in use for the telescope CT1 of HEGRA (see
\ccite{MAGIC:Barrio_Kruger}). In Fig. \ref{fig:MAGICmirrorsandwich} we
can see an exploded view of a mirror element. The reflectivity of
these mirrors, made from an Aluminum alloy, is of the order of
$R(\lambda) > 80\%$ for the wavelength range $320-650\u{nm}$. In
addition, the mirrors are protected by a hard transparent overcoating.

\section{The camera of \MAGIC}

\subsection{General overview}

The camera is the decisive element in the construction of every
Cherenkov telescope. As a detection device, its characteristics will
influence the global performance of the telescope in terms of
efficiency, sensitivity and dynamic range (in the sense of energy
threshold and saturation energy, for instance). Nowadays, every camera
of a Cherenkov telescope is composed of several small detection units
(\emph{pixels}), normally \emph{photo-multipliers} (PMTs).  Therefore,
the overall characteristics of the Cherenkov telescope can be
attributed to:
% 
\begin{enumerate}[a.]
\item The distribution of pixels in the camera (geometry)
\item The pixel size
\item The camera size (field of view --- f.o.v.)
\item The different density  of pixels across  the camera  (related to
  the pixel size)
\item The sensitivity of the pixels
\end{enumerate}

\paragraph{The \emph{distribution of pixels in the camera}} affects only
slightly the performance of the telescope.  While a square matrix
distribution can be easier to build and even more convenient for data
analysis, the most widely distribution used is the \emph{hexagonal
  pattern}. Using hexagonal pixels one can minimize the spaces between
pixels.

\paragraph{The \emph{pixel size}} The first HEGRA camera had a pixel 
size of 0.4\deg, while in the next generation of Cherenkov telescopes
the pixels size will be around 0.10\deg--0.15\deg.  Theoretical
considerations give us an optimal value of $\leq 0.1\deg$. For \MAGIC
we plan to use two different pixel sizes for the camera. In the
central part we will use, in a first step, a pixel size of 0.1\deg.
Then we will have an outer ring of 0.2\deg PMTs. The reason for this
is the following: the low-energy $\gamma$-rays produce showers which
image is closer to the camera center (since they are produce higher in
the atmosphere) and more compact (the distance is bigger, the amount
of light smaller, and the Cherenkov angle in the maximum development
of the shower is smaller than for higher-energy showers). Therefore,
in the center of the camera we need to use a small pixel size. Since
the aberration effects are increasing towards the edge of the camera,
and only the high-energy showers will give trigger even when the image
is far away from the center (for they give much more light), in the
outer part of the image we don't need a very fine pixelization. By
using bigger pixels in the edge, we can lower the already high cost of
the whole camera.

\mirrorsandwichfig

\paragraph{The \emph{camera size} (field of view --- f.o.v.)} is, on
the other  hand,  getting larger. The   benefits of a wider  f.o.v are
mainly:
%
\begin{enumerate}[i)]
\item One can study \emph{extended sources}.
%
\item The reconstruction of high energy showers is more efficient (for
  a small camera we have the so called \emph{edge-effects}: a large
  shower image is truncated at the edge, and therefore the parameters
  of the image and the amount of light are badly estimated, leading to
  a wrong energy reconstruction or, eventually, a rejection of genuine
  $\gamma$-ray events).
%
\item At \emph{high zenith angles} (HZA hereafter) the showers develop
  at much larger distances from the telescope. Therefore, due to the
  scattering and the transverse momentum of the charged particles, the
  \emph{Cherenkov pool}\footnote{The \emph{Cherenkov pool} is the
    region from the axis of the shower till the position of the
    \emph{hump}. Depending on the energy, in this region the density
    of light can increase, decrease or be constant with the distance
    to the shower axis, but can be assumed to be very uniform for
    studies with Cherenkov telescopes.  Beyond the \emph{hump}, the
    density of light drops exponentially.}  is much bigger.  The
  consequence of this is that the showers can trigger the system at
  larger core distances. But this means that the shower image might be
  further from the center than in the \emph{low zenith angle} (LZA)
  case. Therefore, a large camera will be really useful to make use of
  HZA observations.
\end{enumerate}

\paragraph{The \emph{sensitivity of the pixels}} is one of the most
crucial elements in the development of the camera. The performance of
the whole system is directly proportional to the \emph{Quantum
  efficiency} (\QE hereafter) of the pixels.  We will talk about the
different kinds of pixels further below.

\mirrorreflecfig

For \MAGIC we will use a camera size of $\sim$4\deg, with around 600
pixels. %
% (In the simulations we only used a camera of 3.5\deg diameter,
% and all the pixels are the same size, 0.1\deg).  
In Fig.\ref{fig:MAGICcamera} we  can see the geometrical pixel pattern
of  a possible design  of the MAGIC Telescope's  camera (design of the
author, adapted from \ccite{MAGIC:DR}, and actualized).

\QuantumEfffig

\subsection{Light sensors}
%
In the design of \MAGIC we took into account the possibility of an
exchangeable camera, in order to allow modularity and simplicity in
the upgrades.  Several kinds of light sensors have been designed,
improved and tested, and in principle three different camera concepts
are under study:
%
\begin{enumerate}[a.]
\item The \emph{Classical camera}, using standard photo-multipliers
%
\item  The \emph{Standard camera},  which uses hybrid photo-detectors:
  high \QE GaAsP, intensified photocells with avalanche diode readout.
%
\item The \emph{Advanced camera}, a future  solution which employs all
  silicon avalanche photo-diodes.
\end{enumerate}
%
We will locate hollow funnels (\emph{Winston cones}) on top of each
pixel to concentrate the light to the light sensor and to restrict the
acceptance. In addition, wavelength-shifter coatings (WLS) will be
used in order to enhance the detection of UV Cherenkov light by
shifting part of this light into the region where the light sensors
have a greater \QE (see Fig.\ref{fig:QE}).  We will discuss these
different sensors in the following paragraphs.

\MAGICcamerafig

\pixelreadoutfig

\subsubsection{Classical devices: photo-multipliers (PMTs)}
%
In the first stage of the development of the project a classical PMTs
camera will be used. This will affect not only the energy threshold of
the system but also the potential for HZA observations. However, still
the energy threshold is estimated to be around 30\u{GeV}, which is a
significant improvement compared to other existing telescopes or
projects. In the simulations the PMT used for the classical camera is
the EMI 9083-A (see Fig.\ref{fig:QE}) with a peak \QE of about 26\%
and a \emph{mean effective} \QE, \QEeff, of around 17\%, being the
\QEeff defined as the mean value of the \QE in the wavelength range
290--300\u{nm}, folded with the Cherenkov light spectrum, i.e.:
%
\meanQEdefeq
%
where  $I_\mathrm{Ch}(\lambda)$  is the  intensity  of  the \Cherenkov
light emitted at the wavelength $\lambda$.

\subsubsection{Hybrid intensified photo-cells (IPCs)}
%
The long-term aim of the telescope project is to lower the energy
threshold down to 10\u{GeV}.  This goal can be reached by using IPCs
(intensified photo-cells, as they are called by the company Intevac,
one of the developers). These sensors reach more than 40\% \QE in the
region 500--600\u{nm}, with room for more improvements. This, together
with the use of WLS, will increase significantly the \QEeff\ for this
type of camera. By using an avalanche diode in the anode (instead of a
classical diode), a single electron response can be obtained. In
addition, the problem of ion-feedback (ions released in the vicinity
of the anode which accelerated back to the cathode) can be prevented
with an ion deflector.

\subsubsection{Avalanche Photo-Diodes (APDs)}
%
In a future version of the camera of \MAGIC we plan to use all-silicon
Avalanche Photo-Diodes (APDs). These detection devices have a \QEeff\ 
of around $\gtrsim$80\%.  This value is twice the best value achieved
with an IPC using an avalanche photo-diode in the anode. These devices
will allow not only a dramatic increase in the yield of \Cherenkov
photons, but also smaller pixel sizes (0.05\deg--0.07\deg), for they
have active sizes of 3--4\u{mm}. However, the technology is not yet
ready for such application.  The main problem of APDs is the large
noise levels, still unacceptable for our purposes. But the development
continues and in few years it will be possible to use APDs in a \MAGIC
camera.

\bigskip

All camera designs will share a common readout chain, shown in
Fig.\ref{fig:pixelreadout}.

%\MORE%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Other components of the system}
\label{sec:othercomp}

\subsubsection{Frame deformation monitor}
%
The frame of \MAGIC is made out of three layers of carbon fibre tubes.
This construction provides \MAGIC with a very rigid structure.
However, when the telescope moves, there will appear deformations.
Therefore, a system for the monitoring of the deformations in the
frame is needed.  This will provide us with the necessary information
for the study of these deformations depending on several conditions
(orientation of the frame, wind, \ldots). It is planned
\cite{MAGIC:DR} to locate one semitransparent silicon strip sensor at
each corner of the octogonal frame, and two lasers above both sides of
the horizontal (zenital) axis, one for each half of the frame. A
deformation will result in a displacement of the sensors with respect
to the laser beam.  Each sensor can register this displacement and
send it to the central control.

\subsubsection{Tracking monitor}
%
Two systems are foreseen for the monitoring of the position of the
telescope during tracking.  First, two 14bit Gray-coded shaft encoders
will be mounted on both zenital and azimutal axes. The resolution of
these encoders, 0.02\deg, in suficiently smaller than the pixel size.

In addition, a high-sensitivity ($\sim 0.01\u{lux}$), 10\deg field of
view CCD camera will be used to control the position of the camera,
recording its relative position (with the help of some LED markers)
with respect to the stars in the background.  The output of the camera
will be sent to a frame-grabber card on a computer. This allows to
save the information about the position of the camera together with
the event data.

A second CCD will be probably used in order to check the quality of
the sky in the telescope's field of view.

\subsubsection{Active mirror control}
%
Each of the 4-mirror lightweight plate can be oriented individually.
This is usefull for several reasons, the main ones being the
possibility of focus the system at different focal lengths, and the
readjustment of the orientation of the mirrors, even in a
quasi-on-line mode.  For this purpose, an active mirror control system
is required. This system was already outlined on page
\pageref{label:amcs}. In a nutshell, each panel has a small laser
pointer attached in the center, and 3 adjustment screws: one of them
is a double high precision universal joint, that allows free movements
in the XY plane; the other two are driven by stepping motors. One of
the motors is permanently aligned, with no freedom in XY, while the
other can swivel in one direction. The stepping motors are driven via
a multiplexer by a computer-controlled drive circuit.

When adjusting the orientation of one panel, the laser pointer is
turned on, projecting a spot in a white screen in the camera position
(it can be the camera cover in closed position). A CCD video camera
will register the displacement of the laser spot from the reference
center on the screen, and the system will correct consequently the
orientation of the panel. This procedure is repeated for all panels
one by one. The first tests \cite{MAGIC:Wacker_Kruger} indicate that
it will take around 5\u{s} to adjust each panel. Thus a maximum of
about 20\u{min} will be needed for the whole reflector. However, most
mirror panels won't need adjustment: finite element analysis of the
space frame shows that a large part of the central area show no
deformation at all (or at least, if present, it's negligible). This
fact, together with possible pre-saved values of orientations of the
panels, indicates that the whole procedure can be done in a few
minutes.

%An alternative procedure would be to use a distant light source (a
%lamp, a star, \ldots), and move the panel being adjusted, first to one
%side in order to move the spot of light in the camera until it is
%displaced from the center a fixed, given angle (say 1\deg), and then
%to the opposite side, until it is again displaced by the same amount.
%The central position is assumed to correspond with the optimal
%orientation of the panel, and can be interpolated. This procedure is
%free of possible laser pointer misalignment, but cannot be used easily
%for routine corrections.

\subsubsection{Mirror quality monitor}
%
The characteristics of the mirror elements were already discussed in
page \pageref{label:mirrorelement}. An exploded view of one of these
elements is shown in Fig.  \ref{fig:MAGICmirrorsandwich}.

The requirement for the spectral reflectivity if the mirror is
$R(\lambda)>85\%$, with $\lambda$ the wavelength, in the spectral
range of 320--650. In order to prevent the severe degradation of the
mirrors, a continuous monitoring is needed. The reflectivity is one of
the elements that most directly affects the conversion factor
photon-photoelectron of the whole system, and hence it has a severe
influence in the energy determination. Up to now, with the \HEGRA
telescopes, people used a hand-held reflectometer. With this device,
one could check all the mirrors in the telescopes. However, this
procedure with be simply non-viable in \MAGIC: there are too many
mirrors to be monitored by hand, not to speak of the difficult access.
An alternative method is to use a distant deuterium lamp. The light
beam will be chopped and focused onto the telescope.  With the active
mirror control it will be possible to steer the reflected beam of
individual panels onto a photocell or a silicon photodiode, and
measure the reflected light. The use of measurements of different
days, and a reference mirror, will allow to estimate the degradation
of the mirror reflectivity.

\subsubsection{Starfield monitor}
%
Due to the large mirror area of \MAGIC, a considerable amount of light
coming from the starfield will concentrate and form star images in the
camera. As a result of the alt-azimuthal mount of \MAGIC, this image
of the starfield will rotate in the camera plane, when tracking a
object during the night. There are two different approaches to handle
this situation: one can put out of the trigger logic the \emph{hot}
pixels, i.e., the pixels that receive an amount of light from one or
more stars large enough to induce triggers; or, alternatively, one can
try to lower the threshold (to minimize these triggers), and register
the conditions in the camera (and in the f.o.v.), by storing the
positions of the stars in the f.o.v. onto tape, in order to
deconvolute their effect in the data \emph{a posteriori}. For this
purpose, an starfield monitor is needed.

In addition, the starfield monitor could give an input to the tracking
monitoring system and, eventually, to an autoguiding system.

\subsubsection{Observational conditions monitoring}
%
There are many sources of variability in the observational conditions:
not only changes in the weather, but also the orientation of the
telescope, variations in the aerosol content, amount of light of the
night sky, etc.  These elements affect, among other parameters, the
atmospheric transmission. In order to use observed data from different
nights (or even in one night), we need to control these variations.

The atmospheric transmission can be monitorized in many ways:
\begin{itemize}
\item As written above, a CCD will be use in the tracking monitoring.
  A second CCD can be used to check the quality of the sky, since a
  good amount of stars will appear in the telescope's field of view.

\item A separate small optical telescope will be used to check for
  optical variations of known flaring or pulsating $\gamma$-ray
  sources. This telescope can be used for monitoring the brightness
  variations in the region of the sky where \MAGIC is pointing to.

\item Finally, a LIDAR system is planned to be used for recording the
  backscatter profile of the atmosphere.
\end{itemize}

%\section{Data Acquisition}
%

%\section{Observational Programs}
%

%\section{The Telescope CT1 of the HEGRA Collaboration}
%%

\endinput
%
%% Local Variables:
%% mode:latex
%% TeX-master: t
%% End:

%%EOF
